
```{r}
data.name = "mouse_hema"
load(file.path("mouse_hema_tsne_random_p15_pc5_knnMat.RData")) #change as needed  

# Perplexity setting
GCP = 100 #adjust as needed
```

#calculate TSNE scDEED ratio
```{r}
set.seed(123)

# Create a Seurat object
new_seurat_object <- CreateSeuratObject(scale.data)
LayerData(new_seurat_object, layer = 'scale.data', assay = 'RNA') <- scale.data
VariableFeatures(new_seurat_object) <- Features(new_seurat_object)
new_seurat_object <- RunPCA(new_seurat_object)

# Extract t-SNE embeddings
tsne_embeddings <- Y
colnames(tsne_embeddings) <- c("tSNE_1", "tSNE_2")

# Ensure the t-SNE result has proper row names
seurat_cells <- colnames(new_seurat_object)
rownames(tsne_embeddings) <- seurat_cells

# Create a DimReduc object for t-SNE and add it to the Seurat object
new_seurat_object[['tsne']] <- CreateDimReducObject(embeddings = tsne_embeddings, key = 'tSNE_', assay = 'RNA')

# Use t-SNE results in scDEED
resu1 <- scDEED(new_seurat_object, K = 50, reduction.method = 'tsne', perplexity = GCP, rerun = FALSE)

# Extract results from scDEED
resu1_full_results <- resu1$full_results

# Extract and convert cell information from full_results
gcells <- lapply(resu1_full_results[, c("dubious_cells", "trustworthy_cells", "intermediate_cells")], function(x) {
  as.numeric(unlist(str_split(x, ",")))
})

# Convert the list to a data.table for further analysis
gcells_dt <- data.table(
  type = rep(names(gcells), sapply(gcells, length)),
  CellId = unname(unlist(gcells))
)

# Get result summary
result_summary <- gcells_dt[, .N, by = type]
print(result_summary)

# Calculate in/out ratio
in_out_ratio <- result_summary[type == "trustworthy_cells", N] / 
                (result_summary[type == "dubious_cells", N] + result_summary[type == "trustworthy_cells", N])
print(paste("In-Out Ratio:", in_out_ratio))

# Plotting: Highlight dubious and trustworthy cells using the t-SNE reduction
# Adjust as necessary based on scDEED results
dubious_cells <- gcells_dt[type == "dubious_cells", CellId]
trustworthy_cells <- gcells_dt[type == "trustworthy_cells", CellId]

# Plotting highlighted cells on t-SNE
plot <- DimPlot(new_seurat_object, reduction = 'tsne', 
                cells.highlight = list('dubious' = dubious_cells, 'trustworthy' = trustworthy_cells)) + 
        scale_color_manual(values = c('gray', 'blue', 'red')) + 
        ggtitle(paste("t-SNE Plot Highlighting Cells with Perplexity =", GCP))

print(plot)

```

#calculate UMAP scDEED ratio
```{r}
set.seed(123)

# Create a Seurat object
new_seurat_object <- CreateSeuratObject(scale.data)
LayerData(new_seurat_object, layer = 'scale.data', assay = 'RNA') <- scale.data
VariableFeatures(new_seurat_object) <- Features(new_seurat_object)
new_seurat_object <- RunPCA(new_seurat_object)

# Extract UMAP embeddings
umap_embeddings <- as.matrix(Y)
colnames(umap_embeddings) <- c("UMAP_1", "UMAP_2")

# Ensure the UMAP result has proper row names
seurat_cells <- colnames(new_seurat_object)
rownames(umap_embeddings) <- seurat_cells

# Create a DimReduc object for UMAP and add it to the Seurat object
new_seurat_object[['umap']] <- CreateDimReducObject(embeddings = umap_embeddings, key = 'UMAP_', assay = 'RNA')

# Use UMAP results in scDEED
resu1 <- scDEED(new_seurat_object, K = 50, reduction.method = 'umap', perplexity = GCP, rerun = FALSE)

# Extract results from scDEED
resu1_full_results <- resu1$full_results

# Extract and convert cell information from full_results
gcells <- lapply(resu1_full_results[, c("dubious_cells", "trustworthy_cells", "intermediate_cells")], function(x) {
  as.numeric(unlist(str_split(x, ",")))
})

# Convert the list to a data.table for further analysis
gcells_dt <- data.table(
  type = rep(names(gcells), sapply(gcells, length)),
  CellId = unname(unlist(gcells))
)

# Get result summary
result_summary <- gcells_dt[, .N, by = type]
print(result_summary)

# Calculate in/out ratio
in_out_ratio <- result_summary[type == "trustworthy_cells", N] / 
                (result_summary[type == "dubious_cells", N] + result_summary[type == "trustworthy_cells", N])
print(paste("In-Out Ratio:", in_out_ratio))

# Plotting: Highlight dubious and trustworthy cells using the UMAP reduction
# Extract dubious and trustworthy cells to highlight in the plot
dubious_cells <- gcells_dt[type == "dubious_cells", CellId]
trustworthy_cells <- gcells_dt[type == "trustworthy_cells", CellId]

# Plotting highlighted cells on UMAP
plot <- DimPlot(new_seurat_object, reduction = 'umap', 
                cells.highlight = list('dubious' = dubious_cells, 'trustworthy' = trustworthy_cells)) + 
        scale_color_manual(values = c('gray', 'blue', 'red')) + 
        ggtitle(paste("UMAP Plot Highlighting Cells with GCP =", GCP))

print(plot)

```

#check the indices for trustworthy and dubious points
```{r}
length(dubious_cells)
length(trustworthy_cells)

print(dubious_cells)
print(trustworthy_cells)
```

#find cell type distribution of scDEED result
```{r}
#find dubious cell type
dubious_string <- resu1[["full_results"]][1,5] #originally 1,4, changed
dubious_vector <- strsplit(dubious_string, ",")[[1]]
dubious_vector <- as.numeric(dubious_vector)
print(dubious_vector)

dubious_cell_type <- cls[dubious_vector]
print(dubious_cell_type)
dubious_cell_type_counts <- as.data.frame(table(dubious_cell_type))
print(dubious_cell_type_counts)

ggplot(data = dubious_cell_type_counts, aes(x = dubious_cell_type, y = Freq, fill = dubious_cell_type)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Freq), vjust = -0.3, color = "black", size = 3.5) + 
  theme_minimal() +
  labs(title = "Frequency of Dubious Cell Types", x = "Cell Type", y = "Frequency") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 




#find trustworthy cell type
trustworthy_string <- resu1[["full_results"]][1,4] #originally 1,5, changed
trustworthy_vector <- strsplit(trustworthy_string, ",")[[1]]
trustworthy_vector <- as.numeric(trustworthy_vector)
print(trustworthy_vector)

trustworthy_cell_type <- cls[trustworthy_vector]
print(trustworthy_cell_type)
trustworthy_cell_type_counts <- as.data.frame(table(trustworthy_cell_type))
print(trustworthy_cell_type_counts)

ggplot(data = trustworthy_cell_type_counts, aes(x = trustworthy_cell_type, y = Freq, fill = trustworthy_cell_type)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Freq), vjust = -0.3, color = "black", size = 3.5) + 
  theme_minimal() +
  labs(title = "Frequency of Trustworthy Cell Types", x = "Cell Type", y = "Frequency") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

```

#scDEED ratio barplot for cell type
```{r}
ratios <- trustworthy_cell_type_counts$Freq / (trustworthy_cell_type_counts$Freq + dubious_cell_type_counts$Freq)
ratios
ratio_data_frame <- data.frame(
  cell_type = trustworthy_cell_type_counts$trustworthy_cell_type,
  ratio = ratios,
  trustworthy = trustworthy_cell_type_counts$Freq,
  dubious = dubious_cell_type_counts$Freq
)
ratio_data_frame
```

#barplot for scDEED ratio per cell type
```{r}
# Load necessary library
library(ggplot2)

# Order the data frame by ratio in descending order
ratio_data_frame2 <- ratio_data_frame[order(ratio_data_frame$ratio, decreasing = TRUE),]
print(ratio_data_frame2)


# Calculate expanded limits based on the ratio values
expanded_min <- min(ratio_data_frame$ratio, na.rm = TRUE) * 0.9  # Adding some buffer by reducing the minimum by 10%
expanded_max <- max(ratio_data_frame$ratio, na.rm = TRUE) * 1.1  # Increasing the maximum by 10%

# Create the bar plot using coord_cartesian to adjust the y-axis view
ggplot(ratio_data_frame2, aes(x = reorder(cell_type, -ratio), y = ratio, fill = cell_type)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = sprintf("%.3f", ratio)), vjust = -0.3, color = "black", size = 3.5) +
  theme_minimal() +
  labs(title = "Bar Plot of Cell Type Ratios", x = "Cell Type", y = "Ratio") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_cartesian(ylim = c(expanded_min, expanded_max))


```

# Murine Intestinal Data ONLY: use entropy category to do barplot for scDEED ratio
```{r}
entropy_indices <- read.csv("intestinal_processed_entropy_values.csv", header = TRUE, stringsAsFactors = FALSE)

# Ensure the category column is treated as a factor
entropy_indices$category <- as.factor(entropy_indices$category)

# Get indices for each category
category_indices <- split(seq_len(nrow(entropy_indices)), entropy_indices$category)

# Print indices for each category
print(category_indices)

boxplot_data <- data.frame(
  local_stability = numeric(),
  category = factor()
)
```
```{r}
# Calculate the ratio for each category
ratios <- numeric(length(category_indices))  # To store ratios
trustworthy_counts <- numeric(length(category_indices))  # Count trustworthy
dubious_counts <- numeric(length(category_indices))  # Count dubious

# Loop through each category and calculate ratios
for (i in seq_along(category_indices)) {
  indices <- category_indices[[i]]  # Get indices for the category
  
  # Count trustworthy and dubious
  trustworthy_in_category <- sum(indices %in% trustworthy_cells)
  dubious_in_category <- sum(indices %in% dubious_cells)
  
  trustworthy_counts[i] <- trustworthy_in_category
  dubious_counts[i] <- dubious_in_category
  
  # Calculate ratio
  total_in_category <- trustworthy_in_category + dubious_in_category
  ratios[i] <- ifelse(total_in_category > 0, trustworthy_in_category / total_in_category, NA)
}

# Prepare data frame for plotting
barplot_data <- data.frame(
  category = as.factor(names(category_indices)),  # Use category names
  ratio = ratios,
  trustworthy = trustworthy_counts,
  dubious = dubious_counts
)

# Plot the barplot
library(ggplot2)
ggplot(barplot_data, aes(x = category, y = ratio, fill = category)) +
  geom_bar(stat = "identity") +
  #geom_text(aes(label = sprintf("%.2f", ratio)), vjust = -0.5, size = 5) +
  labs(
    title = "Trustworthy Ratio by Category",
    x = "Entropy Category",
    y = "Ratio (#Trustworthy / (#Dubious + #Trustworthy))"
  ) +
  theme_minimal()

```




#check the distribution of dubious and trustworthy cells: 
# 1. specifically for embryoid body umap random GCP=2000, check stability distribution of dubious cell
```{r}
#reload the data first and redo the data preprocessing
#load(file.path("mouse_hema_tsne_random_p15_pc5_knnMat.RData")) #change as needed  

stab.s = 0
k = 50

Y_combined <- data.frame(Y, cls)

# Subset to keep dubious cell
Y_combined <- Y_combined[dubious_vector,]
Y <- Y_combined
cls <- Y_combined$cls

for (i in 1:30) {
  knn.mat[[i]] <- knn.mat[[i]][dubious_vector, ]
}

data.denoise <- data.denoise[dubious_vector, ]

dim1_num <- dim(knn.mat[[1]])[1]  
dim2_num <- dim(knn.mat[[1]])[2]  

for (j in 1:length(knn.mat)) { 
  for (k in 1:dim1_num) {  
    temp_list <- c()  
    temp_count <- 0   
    
    for (x in 1:dim2_num) { 
      if (knn.mat[[j]][k, x] %in% dubious_vector) {  
        temp_count <- temp_count + 1
        temp_list[temp_count] <- knn.mat[[j]][k, x]
      }
    }
    
    # Fill with NAs if the new row is shorter than dim2_num
    if (length(temp_list) != dim2_num) {
      temp_list <- c(temp_list, rep(NA, dim2_num - length(temp_list)))
    }
    
    temp_list <- as.numeric(temp_list)
    knn.mat[[j]][k, ] <- temp_list
  }
}


N <- 30
temp.out <- matrix(ncol = N, nrow = N)

k = dim(knn.mat[[1]])[2]

 for(i in 1:(N-1)){
   for(j in (i+1):N){
     temp=c()
     for(mn in 1:dim(knn.mat[[1]])[1]){ 
        temp[mn]=length(intersect(knn.mat[[i]][mn,1:k],knn.mat[[j]][mn,1:k]))/k  
     }
     temp.out[i,j]=median(temp)
   }
 }

stab.s_dubious <- mean(temp.out, na.rm = TRUE)
print(stab.s_dubious)


#create a boxplot for stab.s
temp_stab_list_dubious <- c()  
temp_stab_list_count_dubious <- 1  
for (i in 1:dim(temp.out)[1]) {  
  for (j in 1:dim(temp.out)[2]) {  
    if (!is.na(temp.out[i, j])) {
      temp_stab_list_dubious[temp_stab_list_count_dubious] <- temp.out[i, j]
      temp_stab_list_count_dubious <- temp_stab_list_count_dubious + 1  
    }
  }
}
print(temp_stab_list_dubious)
boxplot(temp_stab_list_dubious, 
        main = "Stability Score for Dubious Cell Type", 
        ylab = "Scores")

hist(temp_stab_list_dubious,
     main = "Stability Score Distribution for Dubious Cell Type",
     xlab = "stab.s score",
     ylab = "Frequency")

```
# 2. specifically for embryoid body umap random GCP=2500, check stability distribution of trustworthy cell
```{r}
#reload the data first and redo the data preprocessing
#load(file.path("mouse_hema_tsne_random_p15_pc5_knnMat.RData")) #change as needed  

stab.s = 0
k = 50

Y_combined <- data.frame(Y, cls)
Y_combined <- Y_combined[trustworthy_vector,]
Y <- Y_combined
cls <- Y_combined$cls

for (i in 1:30) {
  knn.mat[[i]] <- knn.mat[[i]][trustworthy_vector, ]
}

data.denoise <- data.denoise[trustworthy_vector, ]

dim1_num <- dim(knn.mat[[1]])[1]  
dim2_num <- dim(knn.mat[[1]])[2]  

for (j in 1:30) { 
  for (k in 1:dim1_num) {  
    temp_list <- c()  
    temp_count <- 0   
    
    for (x in 1:dim2_num) {  
      if (knn.mat[[j]][k, x] %in% trustworthy_vector) {  
        temp_count <- temp_count + 1
        temp_list[temp_count] <- knn.mat[[j]][k, x]
      }
    }
    
    # Fill with NAs if the new row is shorter than dim2_num
    if (length(temp_list) != dim2_num) {
      temp_list <- c(temp_list, rep(NA, dim2_num - length(temp_list)))
    }
    
    temp_list <- as.numeric(temp_list)
    knn.mat[[j]][k, ] <- temp_list
  }
  print(j)
}


N <- 30
temp.out <- matrix(ncol = N, nrow = N)

k = dim(knn.mat[[1]])[2]

 for(i in 1:(N-1)){
   for(j in (i+1):N){
     temp=c()
     for(mn in 1:dim(knn.mat[[1]])[1]){ 
        temp[mn]=length(intersect(knn.mat[[i]][mn,1:k],knn.mat[[j]][mn,1:k]))/k  
     }
     temp.out[i,j]=median(temp)
   }
   print(i)
 }

stab.s_trustworthy <- mean(temp.out, na.rm = TRUE)
print(stab.s_trustworthy)


#create a boxplot for stab.s
temp_stab_list_trustworthy <- c()  
temp_stab_list_count_trustworthy <- 1  
for (i in 1:dim(temp.out)[1]) {  
  for (j in 1:dim(temp.out)[2]) {  
    if (!is.na(temp.out[i, j])) {
      temp_stab_list_trustworthy[temp_stab_list_count_trustworthy] <- temp.out[i, j]
      temp_stab_list_count_trustworthy <- temp_stab_list_count_trustworthy + 1  
    }
  }
}
print(temp_stab_list_trustworthy)
boxplot(temp_stab_list_trustworthy, 
        main = "Stability Score for Trustworthy Cell Type", 
        ylab = "Scores")

hist(temp_stab_list_trustworthy,
     main = "Stability Score Distribution for Trustworthy Cell Type",
     xlab = "stab.s score",
     ylab = "Frequency")

```
# 3. draw histogram to see the overlap of dubious and trustworthy cells
```{r}
#histogram
hist(temp_stab_list_trustworthy, col=rgb(1, 0, 0, 0.5),
     xlim=c(min(c(temp_stab_list_trustworthy, temp_stab_list_dubious)), max(c(temp_stab_list_trustworthy, temp_stab_list_dubious))), 
     main="Stability Score Distribution For ESC and CPs, and Other Cell Types", xlab="Stability Score", 
     ylim=c(0, max(c(hist(temp_stab_list_trustworthy, plot=FALSE)$counts, hist(temp_stab_list_dubious, plot=FALSE)$counts))),
     ylab="Stability Score Frequency")

hist(temp_stab_list_dubious, col=rgb(0, 0, 1, 0.5), breaks=20, add=TRUE)  # Add=TRUE to overlay on the first histogram
legend("topright", legend=c("ESP and CPs", "Other Cell Types"), fill=c(rgb(1, 0, 0, 0.5), rgb(0, 0, 1, 0.5)))

#density plot

# Calculate density for both data sets
density1 <- density(temp_stab_list_trustworthy)
density2 <- density(temp_stab_list_dubious)

# Define the range for the Y-axis
max_density <- max(c(density1$y, density2$y))
new_max_density <- max_density * 1.2  # Increase the maximum Y value by 20%

# Set plotting area to be square
par(pty = "s")

# Plot the first density curve with adjusted Y-axis, suppressing x-axis labels initially
plot(density1, col=rgb(1, 0, 0, 0.5),
     xlim=c(min(c(temp_stab_list_trustworthy, temp_stab_list_dubious)), max(c(temp_stab_list_trustworthy, temp_stab_list_dubious))),
     ylim=c(0, new_max_density),  # Apply new Y-axis limits
     main="Stability Score Density For ESC and CPs, and Other Cell Types",
     xlab="Stability Score",
     ylab="Density",
     lwd=4,
     xaxt='n')  # Suppress default x-axis labels

# Add the second density curve
lines(density2, col=rgb(0, 0, 1, 0.5), lwd=4)

# Add the legend with smaller text size
legend("topright", legend=c("ESP and CPs", "Other Cell Types"),
       col=c(rgb(1, 0, 0, 0.5), rgb(0, 0, 1, 0.5)), lwd=2, cex=0.8)

# Add custom x-axis with increments of 0.2
axis(1, at=seq(0, 1, by=0.2), labels=seq(0, 1, by=0.2))  # Add tick marks and labels at increments of 0.2

# Define x and y limits if needed
x_limits <- range(c(temp_stab_list_trustworthy, temp_stab_list_dubious))
y_max <- max(c(hist(temp_stab_list_trustworthy, plot=FALSE)$counts, 
               hist(temp_stab_list_dubious, plot=FALSE)$counts))


# Histogram of the trustworthy list
hist(temp_stab_list_trustworthy, col=rgb(1, 0, 0, 0.5), xlim=x_limits,
     main="Stability Score Distribution For Dubious Cell and Trustworthy Cell",
     xlab="Stability Score", ylim=c(0, y_max),
     ylab="Stability Score Frequency", breaks=20)

# Overlay histogram of the dubious list
hist(temp_stab_list_dubious, col=rgb(0, 0, 1, 0.5), add=TRUE)

# Add legend
legend("topright", legend=c("Trustworthy Cell", "Dubious Cell"), fill=c(rgb(1, 0, 0, 0.5), rgb(0, 0, 1, 0.5)))

```
# 4. apply two sample t test to dubious and trustworthy cell lists
```{r}
#test whether the two stability lists have equal variance (F test)
# H0: equal variance; Ha: NOT equal variance
var.test(temp_stab_list_trustworthy, temp_stab_list_dubious)

# Welch's t-test. If reject H0, var.equal=FALSE; else var.equal=TRUE
t_test_result <- t.test(temp_stab_list_trustworthy, temp_stab_list_dubious, var.equal = FALSE,conf.level = 0.99)
print(t_test_result)
```
# 5. draw boxplots on the same plot
```{r}

boxplot_data <- data.frame(
  Scores = c(temp_stab_list_trustworthy, temp_stab_list_dubious),
  Group = factor(
    c(rep("Trustworthy Cell", length(temp_stab_list_trustworthy)),
      rep("Dubious Cell", length(temp_stab_list_dubious)))
  )
)

# Load required libraries
library(ggplot2)
library(ggsignif)

# Create the boxplot with significance tests
ggplot(boxplot_data, aes(x = Group, y = Scores, fill = Group)) + 
  geom_boxplot() +
  geom_signif(comparisons = list(c("Trustworthy Cell", "Dubious Cell")),
              map_signif_level = TRUE) +         #if 2 sample t test shows significance, = TRUE, else = FALSE
  labs(title = "Comparison of Stability Scores",
       y = "scDEED Score",
       x = "") + 
  scale_fill_manual(values = c("red", "blue"))


```

# 6. check in/out ratio for ALL data
```{r}
all_ratio <- length(trustworthy_vector) / (length(dubious_vector) +  length(trustworthy_vector))
all_ratio
```


# 7. check ESC and CPs (stable cell type) trustworthy/(dubious+trustworthy) ratio (in/out ratio)
```{r}
#check the indices of ESC and CPs for embryoid body data
esc_indices <- which(cls == "ESC")
cp_indices <- which(cls == "CPs")
esc_cp <- c(esc_indices,cp_indices)

#extract the corresponding points from scDEED result column
#dubious_vector; trustworthy_vector
esc_cp_in_dubious <- intersect(esc_cp,dubious_vector)
esc_cp_in_trustworthy <- intersect(esc_cp,trustworthy_vector)

specific_cell_ratio <-  length(esc_cp_in_trustworthy) / (length(esc_cp_in_dubious) + length(esc_cp_in_trustworthy))
print(length(esc_cp_in_trustworthy))
print(length(esc_cp_in_dubious))
print(specific_cell_ratio)


#extract points that are not ESC/CPs from scDEED result column
cls_list <- as.numeric(seq(1, length(cls)))
other_indices <- cls_list[-esc_cp]
other_in_dubious <- intersect(other_indices,dubious_vector)
other_in_trustworthy <- intersect(other_indices,trustworthy_vector)

other_cell_ratio <-  length(other_in_trustworthy) / (length(other_in_dubious) + length(other_in_trustworthy))
print(length(other_in_trustworthy))
print(length(other_in_dubious))
print(other_cell_ratio)


```




#check the distribution of ESC and CPs(stable cell type) and Other Cells(unstable cell types):
# 1. specifically for embryoid body umap random GCP=2000, average stability for other cell type (unstable)
```{r}
#reload the data first and redo the data preprocessing
#load(file.path("mouse_hema_tsne_random_p15_pc5_knnMat.RData")) #change as needed  

stab.s = 0
k = 50
Y_combined <- data.frame(Y, cls)

# Find stable data rows
specific_cls <- which(Y_combined$cls %in% c("ESC", "CPs"))

# Subset to keep all unstable types by deleting stable types
Y_combined <- subset(Y_combined, !(cls %in% c("ESC", "CPs")))
Y <- Y_combined
cls <- Y_combined$cls

for (i in 1:30) {
  knn.mat[[i]] <- knn.mat[[i]][-specific_cls, ]
}

data.denoise <- data.denoise[-specific_cls, ]

dim1_num <- dim(knn.mat[[1]])[1]  
dim2_num <- dim(knn.mat[[1]])[2]  

for (j in 1:length(knn.mat)) { 
  for (k in 1:dim1_num) {  
    temp_list <- c()  
    temp_count <- 0   
    
    for (x in 1:dim2_num) {  
      if (!(knn.mat[[j]][k, x] %in% specific_cls)) {  
        temp_count <- temp_count + 1
        temp_list[temp_count] <- knn.mat[[j]][k, x]
      }
    }
    
    # Fill with NAs if the new row is shorter than dim2_num
    if (length(temp_list) != dim2_num) {
      temp_list <- c(temp_list, rep(NA, dim2_num - length(temp_list)))
    }
    
    temp_list <- as.numeric(temp_list)
    knn.mat[[j]][k, ] <- temp_list
    
    #print(paste("Matrix", j, "Row", k, "Processed."))
  }
}

N <- 30
temp.out <- matrix(ncol = N, nrow = N)
k = dim(knn.mat[[1]])[2]

for (i in 1:(N-1)) {
  for (j in (i+1):N) {
    temp <- c()
    for (mn in 1:dim(knn.mat[[1]])[1]) {  
      temp[mn] <- length(intersect(knn.mat[[i]][mn, 1:k], knn.mat[[j]][mn, 1:k])) / k  
    }
    temp.out[i, j] <- median(temp)
  }
  print(i)
}

stab.s <- mean(temp.out, na.rm = TRUE)
print(stab.s)

#create a boxplot for stab.s
temp_stab_list <- c()  
temp_stab_list_count <- 1  

for (i in 1:dim(temp.out)[1]) {
  for (j in 1:dim(temp.out)[2]) {
    if (!is.na(temp.out[i, j])) {
      temp_stab_list[temp_stab_list_count] <- temp.out[i, j]
      temp_stab_list_count <- temp_stab_list_count + 1
    }
  }
}

print(temp_stab_list)

boxplot(temp_stab_list,
        main = "Stab.s for Other Unstable Cell Type",
        ylab = "Scores")

hist(temp_stab_list,
     main = "Stability Score Distribution for Other Unstable Cell Type",
     xlab = "stab.s score",
     ylab = "Frequency")

# Store data to try to compare results with previous results
temp_stab_list_2 <- temp_stab_list
cls_2 <- cls
Y_2 <- Y
Y_combined_2 <- Y_combined
data_denoise_2 <- data.denoise
knn.mat_2 <- knn.mat
stab.s_2 <- stab.s
```
# 2. specifically for embryoid body umap random GCP=2500, average stability for 2 cell types (stable)
```{r}
#reload the data first and redo the data preprocessing
#load(file.path("mouse_hema_tsne_random_p15_pc5_knnMat.RData")) #change as needed

stab.s_1 = 0
k = 50
Y_combined <- data.frame(Y, cls)

# Find rows that are not stable cell types
specific_cell_type2 <- which(!(Y_combined$cls %in% c("ESC", "CPs")))

# Subset to keep only stable cell types
Y_combined <- subset(Y_combined, cls %in% c("ESC", "CPs"))
Y <- Y_combined
cls <- Y_combined$cls

# Adjust knn.mat by removing rows corresponding to specific_cell_type2
for (i in 1:30) {
  knn.mat[[i]] <- knn.mat[[i]][-specific_cell_type2, ]
}

data.denoise <- data.denoise[-specific_cell_type2, ]

dim1_num <- dim(knn.mat[[1]])[1]  
dim2_num <- dim(knn.mat[[1]])[2]  

for (j in 1:length(knn.mat)) {  
  for (k in 1:dim1_num) {  
    temp_list <- c()  
    temp_count <- 0   
    
    for (x in 1:dim2_num) {  
      if (!(knn.mat[[j]][k, x] %in% specific_cell_type2)) {  
        temp_count <- temp_count + 1
        temp_list[temp_count] <- knn.mat[[j]][k, x]
      }
    }
    
    # Fill with NAs if the new row is shorter than dim2_num
    if (length(temp_list) != dim2_num) {
      temp_list <- c(temp_list, rep(NA, dim2_num - length(temp_list)))
    }
    
    temp_list <- as.numeric(temp_list)
    knn.mat[[j]][k, ] <- temp_list
    
    #print(paste("Matrix", j, "Row", k, "Processed."))
  }
}

N <- 30
temp.out <- matrix(ncol = N, nrow = N)
k = dim(knn.mat[[1]])[2]

for (i in 1:(N-1)) {
  for (j in (i+1):N) {
    temp <- c()
    for (mn in 1:dim(knn.mat[[1]])[1]) {
      temp[mn] <- length(intersect(knn.mat[[i]][mn, 1:k], knn.mat[[j]][mn, 1:k])) / k
      print(mn)
    }
    temp.out[i, j] <- median(temp)
  }
}

stab.s <- mean(temp.out, na.rm = TRUE)
print(stab.s)

#create a boxplot for stab.s
temp_stab_list <- c()
temp_stab_list_count <- 1

for (i in 1:dim(temp.out)[1]) {
  for (j in 1:dim(temp.out)[2]) {
    if (!is.na(temp.out[i, j])) {
      temp_stab_list[temp_stab_list_count] <- temp.out[i, j]
      temp_stab_list_count <- temp_stab_list_count + 1
    }
  }
}

print(temp_stab_list)

boxplot(temp_stab_list,
        main = "Boxplot of stab.s for Stable Cell Type",
        ylab = "Values")

hist(temp_stab_list,
     main = "Stability Score Distribution for Stable Cell Type",
     xlab = "stab.s score",
     ylab = "Frequency")

```
# 3. draw histogram to see the overlap of stable and unstable cells
```{r}
#histogram
hist(temp_stab_list, col=rgb(1, 0, 0, 0.5),
     xlim=c(min(c(temp_stab_list, temp_stab_list_2)), max(c(temp_stab_list, temp_stab_list_2))), 
     main="Stability Score Distribution For ESC and CPs, and Other Cell Types", xlab="Stability Score", 
     ylim=c(0, max(c(hist(temp_stab_list, plot=FALSE)$counts, hist(temp_stab_list_2, plot=FALSE)$counts))),
     ylab="Stability Score Frequency")

hist(temp_stab_list_2, col=rgb(0, 0, 1, 0.5), breaks=20, add=TRUE)  # Add=TRUE to overlay on the first histogram
legend("topright", legend=c("ESP and CPs", "Other Cell Types"), fill=c(rgb(1, 0, 0, 0.5), rgb(0, 0, 1, 0.5)))

#density plot

# Calculate density for both data sets
density1 <- density(temp_stab_list)
density2 <- density(temp_stab_list_2)

# Set plotting area to be square
par(pty = "s")

# Plot the first density curve, suppressing the default x-axis labels
plot(density1, col=rgb(1, 0, 0, 0.5), 
     xlim=c(min(c(temp_stab_list, temp_stab_list_2)), max(c(temp_stab_list, temp_stab_list_2))),
     main="Stability Score Density for ESC and CPs, and Other Cell Types", 
     xlab="Stability Score", 
     ylab="Density", 
     lwd=2,
     xaxt='n')  # Suppress default x-axis labels

# Add the second density curve
lines(density2, col=rgb(0, 0, 1, 0.5), lwd=2)

# Add custom x-axis with increments of 0.2
axis(1, at=seq(0, 1, by=0.2), labels=seq(0, 1, by=0.2))

# Add the legend
legend("topright", legend=c("ESP and CPs", "Other Cell Types"), 
       col=c(rgb(1, 0, 0, 0.5), rgb(0, 0, 1, 0.5)), lwd=2, cex=0.7)

# Define x and y limits for histogram
x_limits <- range(c(temp_stab_list, temp_stab_list_2))
y_max <- max(c(hist(temp_stab_list, plot=FALSE)$counts, 
               hist(temp_stab_list_2, plot=FALSE)$counts))

# Plot the histogram for the trustworthy list, suppressing default x-axis
hist(temp_stab_list, col=rgb(1, 0, 0, 0.5), xlim=x_limits,
     main="Stability Score Distribution for Unstable Cell and Stable Cell",
     xlab="Stability Score", ylim=c(0, y_max),
     ylab="Stability Score Frequency", breaks=20,
     xaxt='n')  # Suppress default x-axis labels

# Overlay histogram of the dubious list
hist(temp_stab_list_2, col=rgb(0, 0, 1, 0.5), add=TRUE)

# Add custom x-axis with increments of 0.2
axis(1, at=seq(0, 1, by=0.2), labels=seq(0, 1, by=0.2))

# Add legend for the histogram
legend("topright", legend=c("Unstable Cell", "Stable Cell"), fill=c(rgb(1, 0, 0, 0.5), rgb(0, 0, 1, 0.5)), cex=0.8)

```
# 4. apply two sample t test to unstable and stable cell lists
```{r}
#test whether the two stability lists have equal variance (F test)
# H0: equal variance; Ha: NOT equal variance
var_test_result <- var.test(temp_stab_list, temp_stab_list_2) 
print(var_test_result)

# Welch's t-test
# Welch's t-test. If reject H0, var.equal=FALSE; else var.equal=TRUE
t_test_result <- t.test(temp_stab_list, temp_stab_list_2, var.equal = FALSE, conf.level = 0.99)
print(t_test_result)

```
# 5. draw boxplots on the same plot
```{r}
boxplot_data <- data.frame(
  Scores = c(temp_stab_list, temp_stab_list_2),
  Group = factor(
    c(rep("Ratio = 1 Cell", length(temp_stab_list)),
      rep("Other Cell", length(temp_stab_list_2)))
  )
)

# Create the boxplot with significance tests
ggplot(boxplot_data, aes(x = Group, y = Scores, fill = Group)) +
  geom_boxplot() +
  geom_signif(comparisons = list(c("Ratio = 1 Cell", "Other Cell")),
              map_signif_level = TRUE) +
  labs(title = "Comparison of Stability Scores",
       y = "scDEED Score",
       x = "") +
  scale_fill_manual(values = c("blue", "red"))

```
#find average stability for ESC+CPs and other cell
```{r}
print(mean(temp_stab_list))
print(mean(temp_stab_list_2))
```


